---
title: "Classification and Regression Analysis using British Seatbelt Data"
output: html_notebook
---

## Where to find datasets?: The "datasets" library
```{r}
library(datasets)

seatbelts <- datasets::Seatbelts

df <- as.data.frame(seatbelts)
# df$Date <- rownames(seatbelts)

print(head(df))
```



# Load time series modeling library (XTS and Zoo) and using them to decode the data column in the "datasets::Seatbelts" dataset
```{r}
library(zoo)
library(xts)
# df$Date <- as.yearmon(time(seatbelts))
df$Date <- as.Date(time(seatbelts))

df$law <- as.factor(df$law)
sapply(df,class)
```

## Exploratory analysis of the difference between classes of law = 1 and law = 0 w.r.t. DriversKilled and VanKilled
```{r}
boxplot(DriversKilled ~ law, data = df)
t.test(DriversKilled ~ law, data = df) #Establish the statistical significance of the difference of means between law = 1 and law = 0 w.r.t. DriversKilled

boxplot(VanKilled ~ law, data = df)
t.test(VanKilled ~ law, data = df) #Establish the statistical significance of the difference of means between law = 1 and law = 0 w.r.t. VanKilled
```

##  Logistic Regression. Use "binomial" to set it as categorical data
```{r}
model <- glm (formula = law ~ DriversKilled + VanKilled, family = "binomial", data = df)
summary(model)

exp(model$coefficients) # 
```

## Create predictions of the fit logistic model against our original data - each row being an independent evalutation of the model
```{r}
predictionsOfLawInEffect <- predict(model, newdata = df)
# Categorize the predicted 
predictionsOfLawInEffect_CATEGORICAL <- predictionsOfLawInEffect > -2
```


```{r}

print(table(df$law, predictionsOfLawInEffect_CATEGORICAL)) #Confusion matrix
```

## Visualize a probability density plot of the predicted odds of the response of law = 1 in the predicted set of evaluated records 

# ODDS = p / (1 - p); where ODDS = exp(predictionsOfLawInEffect)
# => (1 - p)*ODDS = p
# => p = ODDS/(1+ODDS)
```{r}
plot(density(exp(predictionsOfLawInEffect))) 
plot(density(exp(predictionsOfLawInEffect)/(1+exp(predictionsOfLawInEffect))))  # plot of the "ODDS" of each row being classified for the law being in effect
print(summary(exp(predictionsOfLawInEffect)/(1+exp(predictionsOfLawInEffect))))
```


## Boxplot to find the optimal probability cutoff for dichotimization of the law = 1 status
```{r}
boxplot(exp(predictionsOfLawInEffect)/(1+exp(predictionsOfLawInEffect)) ~ df$law)
```

## Perform Linear Regression as an example use-case to predicted drivers that died as a function of the law status being in effect
```{r}
# Linear Regression for REGRESSION of the number of people that died ON the Xs
#  /regressors law-being-in-effect and total number of drivers
# Y = number of people that died
# X = [law, drivers, PetrolPrice]

model2 <- lm (DriversKilled ~ law, data = df)
summary (model2)

```

## Export df to a file:
```{r}
write.csv(df, row.names = F, file = "britishSeatBeltStudy.csv")
```